Supplementary methods

### 2020-08-31 text/supplemental_methods.Rmd
place figures in supplementary methods

Time is split into two components. $A$ going from 0 up to the age of the patient
at diagnosis and then remaining fixed, whereas $t$ starts at 0 upon diagnosis
and represents the days since diagnosis, a field recorded in the dataset. $D(t)$
is an indicator for whether diagnosis has occurred or not (before diagnosis
$D(t) = 0$, after $D(t) = 1$). The model assumes that the risk of death is
$R_{20}$ at age 20 years and increases exponentially with a constant rate $a$ up
to the age at diagnosis. Risk of death assumes a new rate $ka$ thereafter.
Though the model extends to earlier ages, relatively few patients were under age
20.

@@

The product $k_{tis}k_{cl}$ is obviously symmetric between the two factors. To
disambiguate the meaning of the two, the prior distribution of $k_{cl}$ has peak
probability at 1 and a relatively narrow width, due to the fact that it is a
tissue specific multiplicative modifying factor. This allows the overall
magnitude of $k$ to be primarily determined by the $k_{tis}$ value whose prior
is significantly less constrained so that there is a wider range of risk across
tissue types.

**Supplementary figure SM2** gives a comparison between the post-fit
survival curves for men and women vs the CDC data it is based on. This
comparison does not distinguish on ethnicity for simplicity in
visualization.

**Supplementary Figure SM2** Comparison between CDC data and ODE model fit for
males and females of European descent.

>>>>

Time is split into two components. $A$ going from 0 up to the age of the patient
at diagnosis and then remaining fixed, whereas $t$ starts at 0 upon diagnosis
and represents the days since diagnosis, a field recorded in the dataset. $D(t)$
is an indicator for whether diagnosis has occurred or not (before diagnosis
$D(t) = 0$, after $D(t) = 1$). The model assumes that the risk of death is
$R_{20}$ at age 20 years and increases exponentially with a constant rate $a$ up
to the age at diagnosis. Risk of death assumes a new rate $ka$ thereafter.
Though the model extends to earlier ages, relatively few patients were under age
20 (**supplementary figure SM2**).

**Supplementary Figure SM2** Comparison between CDC data and ODE model fit for
males and females of European descent. This comparison does not distinguish on
ethnicity for simplicity in visualization.

@@

The product $k_{tis}k_{cl}$ is obviously symmetric between the two factors. To
disambiguate the meaning of the two, the prior distribution of $k_{cl}$ has peak
probability at 1 and a relatively narrow width, due to the fact that it is a
tissue specific multiplicative modifying factor. This allows the overall
magnitude of $k$ to be primarily determined by the $k_{tis}$ value whose prior
is significantly less constrained so that there is a wider range of risk across
tissue types (**supplementary figure SM3**).
